# Week 1 Discussion: tmap

## Load packages

```{r}
#| eval: true
#| message: false
#| warning: false

library(here)
library(tidyverse)
library(sf)
library(stars)
library(tmap)
```

## Import data

```{r}
ei_points <- st_read(here("data", "week1", "ei_points.gpkg")) |> # st_read is for vector data
  filter(type == "volcano")
ei_elev <- read_stars(here("data", "week1", "ei_elev.tif")) # read_stars is for raster data
ei_borders <- st_read(here("data", "week1", "ei_border.gpkg"))
ei_roads <- st_read(here("data", "week1", "ei_roads.gpkg"))

# Discrete things with edges are vectors (road, coastline, volcano). A value that exists everywhere (elevation) is a raster.
```

## Create a single-object map with tmap
- Create a map of the outline Rapa Nui using `tm_shape()` + `tm_borders()`

```{r}
tm_shape(ei_borders) +
    tm_borders() +
    tm_title('Rapa Nui')
```

## Create a multiple-object map with tmap
Expand on your single-object map to create a map of Rapa Nui and…

…denote the island’s borders and continuous elevation
…denote the island’s volcanoes and roads
…play with the color palette and essential map elements

```{r}
tmap_mode('view')

ei_volcano_map <- tm_shape(ei_elev) +
    tm_graticules() + # adds lat and long lines
    tm_raster(col.scale = tm_scale_continuous(values = "-rd_yl_gn"), 
            col.legend = tm_legend("Elevation (m asl)")) +
    tm_shape(ei_borders) +
        tm_borders() +
    tm_shape(ei_roads) +
        tm_lines() +
    tm_shape(ei_points) +
        tm_symbols(shape = 24, # adds triangle-shaped symbols
                size = 'elevation', # size is based on elevation
                size.legend = tm_legend('Volcanoes (m asl)')) +
    tm_compass(position = c('right', 'top')) +
    tm_scalebar() + # adds scale bar to map
    tm_title('Volcanoes on Rapa Nui')
    
ei_volcano_map
```

## Create an interactive map with tmap
Switch from the default "plot" mode to "view"

See above. Called `tmap_mode("view")`

## Create a multiple-object map with ggplot2
As a bonus challenge, reproduce your multiple-object map from earlier using `ggplot2`!

Install ggspatial package
```{r}
install.packages('ggspatial')
```

Load packages
```{r}
library(ggplot2)
library(ggspatial)
```

Example from [course website](https://eds-223-geospatial.github.io/course-materials/discussions/week1-discussion.html)
```{r}
# extract lat & long from geom column 
volcanoes_point <- ei_points %>%
  mutate(lon = unlist(map(ei_points$geom,1)), # longitude = first component (x)
         lat = unlist(map(ei_points$geom,2))) # latitude = second component (y)

ggplot() +
  # first add Easter Isand borders
  geom_sf(data = ei_borders, color = "#212529") +
  # add elevation raster
  geom_stars(data = ei_elev) +
  scale_fill_distiller(name = "Elevation (m asl)", # legend title with units
                       palette = "RdYlGn",         # red-yellow-green palette (no need to reverse)
                       na.value = "lightblue") +  # set NA color (background ocean)
  # add road vector
  geom_sf(data = ei_roads, color = "#343a40") +
  # add volcano vector
  geom_point(data = volcanoes_point, 
             aes(x = lon, y = lat, 
                 size = elevation), # point size depends on attribute value
             shape = 17,            # triangle
             color = "#22577a") +
  scale_size_continuous(name = "Volcanoes (m asl)") +       # legend title with units
  ggspatial::annotation_north_arrow(location = "br",        # bottom right
                                    which_north = "true") + # point to north pole
  ggspatial::annotation_scale(location = "bl",              # bottom left
                              width_hint = 0.5) +           # proportion of map area the scalebar should occupy
  labs(title = "Rapa Nui") +
  theme_minimal()

```